---
title: "loops_hw_warren_cook"
author: "Kelsey Warren & Taylor Cook"
date: "2025-03-06"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE)
knitr::opts_chunk$set(error=TRUE)

library(tidyverse)
library(purrr)
library(ggpubr)
```

# Homework highway

4 car types

Imagine with have 4 different car types - and we know how often each
occurs:

-   car A mass 20000 kg, area 22 m^2^

-   car B mass 65000 kg, area 30 m^2^

-   car C mass 38000 kg area 22 m^2^

-   car D mass 5000 kg area 10 m^2^

Mean highway speed is 100 km/hr

```{r sampling2, eval=TRUE, echo=TRUE}

# source("../R/autopower.R")
source("~/Desktop/262 Computing/R/automobile_power_generation.R")
# generate a structure to store info on our possible cars
possible_cars = data.frame(name = c("A","B","C","D"),mass=c(10000,65000,38000,5000), area = c(22,30,22,10))

# first look at how results vary for mean speed say 100km/hr
# do conversion
# aka converting km/hr to mi/hr or opposite
speed_base = 80 * 0.28

# because I have one mass and area for each car and only 1 speed
# I can estimate power for each car type
# add to the data structure
# R is running a for loop behind the scenes for us as it calculates power for each car in the data frame
possible_cars$power = automobile_power_generation(V=speed_base, A = possible_cars$area, m=possible_cars$mass)

head(possible_cars)

# show results                         
ggplot(possible_cars, aes(x=mass, y=power, fill=as.factor(area)))+geom_col()+labs(y="Power W", x="Mass (kg)", fill="Area (m2)")
  

```

# Building a highway

What could be the total power consumed if there are 200 cars on this
highway each hour, they are travelling at a range of speeds - mean is
100km/hr and speeds tend to be log-normally distributed)

How would the total power consumed vary by car So if all cars are car A;
OR all cars are car B OR all cars are car C

We will use *sample* here to generate speeds for our 100 cars and look
at different ways to repeat power calculation for different cars

-   repeating by hand
-   *for* for repetition - a more standard way of looping available in
    many langugaes
-   *pmap* for repetition - a efficient way of looping in R
    -   pmap hides the loop from us, part of the purr package

# First lets do it 'by hand'

```{r byhand}

# what is I want to estimate average power use given  each car

possible_cars

# use sample to generate a distribution of speeds

# assume a log normal distribution of speeds with mean 100km/hr, and standard deviation that is 10% of the mean

# recall our function needs speed in m/s not km/hr so we will also do a conversion
# lets get a sample of a 200 speeds- we could also do this by actually measuring speeds

nsample = 200
mean_speed = (80*0.277)

speeds = rnorm(mean=mean_speed, sd=(10*0.277), nsample)
summary(speeds)

plot(density(speeds), ylab="Distribution of Speeds in (m/s)")


# how do we run each car for all speeds 

# first lets to it by hand for the first car - the first row in our possible cars matrix
possible_cars[1,]

# we could do it by hand
# below examples are saying if all highway cars are same mass/area as car A, let's see the powers for all the speeds (then next chunk for B, next chunk for C)
powerA = automobile_power_generation(V=speeds, A = possible_cars$area[1], m=possible_cars$mass[1])
# lets look at what we get
summary(powerA)

# next car (row 2)
powerB = automobile_power_generation(V=speeds, A = possible_cars$area[2], m=possible_cars$mass[2])
# lets look at what we get
summary(powerB)

# next car (row 3)
powerC = automobile_power_generation(V=speeds, A = possible_cars$area[3], m=possible_cars$mass[3])
# lets look at what we get
summary(powerC)

# next car (row 4)
powerD = automobile_power_generation(V=speeds, A = possible_cars$area[4], m=possible_cars$mass[4])
# lets look at what we get
summary(powerD)

summary(speeds)


# we could put this together
powerall1 = cbind.data.frame(powerA, powerB, powerC, powerD)
colnames(powerall1)=possible_cars$name


# for plotting sometimes its useful to turn columns in to rows - we can use an R function
# called pivot_longer (part of the tidyverse package) to do this
# compare powerall1 and powerallr1 to see what pivot_longer does
powerallr1 = powerall1 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerallr1)

# quick visualization
# lets save it so that we can compare
method1_plot = ggplot(powerallr1, aes(car,power, fill=car))+geom_boxplot()+ggtitle("By Hand")
method1_plot



```

# Second- classic looping *for*

Doing this by hand would be hard if we had many different cars - can we
automate?

YES

For loop

Lets do this one more time using a *for* loop

```{r withforloop}

# the first part, generating speeds is the same
# what is I want to estimate average power use given  each car

possible_cars


# how do we run each car for all speeds 
# we use a for loop to cycle through
# we need to create a data frame to store results - as above
# one column for each car and one row for each speed


# create the data frame before using the for loop to fill it
powerall2 = as.data.frame(matrix(nrow=length(speeds), ncol=nrow(possible_cars)))
# because we don't initialize it - values are NA
head(powerall2)

# more checking the data frame
# how many cars area there
nrow(possible_cars)
ncol(powerall2)
# for loops use an index - in this case "i" but you could use anything - it repeats
# anything between the {} for each values of i between 1 and nrow(possible_car) (which is 3 in our case)

# index in to a matrix (like powerall3) is by row and column powerall3[2,5] is 2nd row and 5th column



# FOR LOOP (sometimes for loops take a long time to run)
for (i in 1:ncol(powerall2)) {
  powerall2[,i] = automobile_power_generation(A=possible_cars$area[i], m=possible_cars$mass[i], V=speeds)
}



# now it looks like above
head(powerall2)
# we can add column names
colnames(powerall2)=possible_cars$name

# plotting is the same as above

# for plotting sometimes its useful to turn columns in to rows
powerallr2 = powerall2 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerallr2)

# quick visualization
method2_plot = ggplot(powerallr2, aes(car,power, fill=car))+geom_boxplot()+ggtitle("For Loop")

ggarrange(method1_plot, method2_plot, nrow=2)

powerall1 %>% map(mean)
powerall2 %>% map(mean)

```

# Third using R built in tools

*pmap* works quickly but it is unique to R Other programming language
(and R) use what are called loops - where repetition is more explicit

*pmap* is available in the *purr* library

*mapply* is another R option

*pmap* - runs a function for a list of parameters - using other
parameters for each iteration

```{r withpmap}



# how do we run each car for all speeds 
# pmap runs a function for each value in a list of parameters, with other parameters set for each iteration


powerall3 = pmap(list(A = possible_cars$area, m=possible_cars$mass), automobile_power_generation, V=speeds)

# lets turn to a data frame for easier graphing
# we can add column names
powerall3 = as.data.frame(powerall2, col.names=possible_cars$name)

head(powerall3)


# for plotting sometimes its useful to turn columns in to rows
powerallr3 = powerall3 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerallr2)

# quick visualization

method3_plot = ggplot(powerallr3, aes(car,power, fill=car))+geom_boxplot()+ggtitle("pmap")
method3_plot

# put plots side by side
# to confirm that they look similar
ggarrange(method1_plot, method2_plot, method3_plot, nrow=3)

# compare values
head(powerallr2)
head(powerallr1)
head(powerallr3)


```

# Assignment 4

For this assignment you will write a new Quartro document based on the
analysis of power required for different *possible_cars* from
looping.RMD; You will

-   Add an additional super light car with mass 5,000 kg and a surface
    area 10m2

-   Compute the mean and maximum power for Cars A, B, C and for your new
    car, assuming that average highway speed is 80 km/hr with a standard
    deviation of 10km/hr; Use 2 different methods to compute the mean
    and maximum power for each car type; a) **FOR** loop and b) the
    **pmap** function from **purr**

-   Create two boxplots (one for each method (**FOR** and **pmap**))
    that show the range of power consumption (across sampled speeds) for
    each car type.

Put the Quatro in your assignment github and submit a link on Canvas for
Assignment 4 when its ready to be graded

**Grading Rubric**

1.  Correct R code for adding the super light car (5pts)
2.  Values for mean and maximum power for all 4 cars are in expected
    ranges (remember you are sampling so values will differ slightly)
    (5pts)
3.  Used both methods (**FOR** and **pmap**) correctly (4pts)
4.  Boxplots (6pts)

<!-- -->

a)  correctly show range of power consumption for each car type for the
    two methdos
b)  good presentation (e.g axis labels, colors)

<!-- -->

5.  Good coding practice (e.g Adequate documentation throughout) (5pts)

Total out of 25pts
